Combinatorics of the geometry of Wilson loop diagrams

### 2019-02-26 General WLD combinatorics.tex
General WLD combinatorics.tex

This in turn defines two sets of propagators, $\{\cP_{in}, \cP_{out}\}$, where (for $\bullet \in \{in, \; out\}$) \bas \cP_{\bullet} = \{ q \in (\cP \setminus p) | V(q) \cap V_{\bullet}(p)  \neq \emptyset \} \;.\eas The set $\cP_{in}$ (resp. $\cP_{out}$) is the set of propagators, not equal to $p$ \hlfix{whose support contains points in $V_{in}(p, <_*)$ (resp. $V_{in}(p, <_*)$), where $<_*$ refers to the appropriate ordering.}{Note, this means that paralell propagators are neither on the inside and outside of $p$. Is that okay?} If $W$ has no propagators crossing $p$ and no other propagator supported entirely on $V(p)$, then the set $\{\cP_{in}, p, \cP_{out}\}$ partitions $\cP$. Under these circumstances, the set $\cP_{in}$ defines the propagators with end points on one side of $p$, while $\cP_{out}$ defines the propagators with end points on the other side. In particular, for (weakly) admissible Wilson loop diagrams, the set $\{\cP_{in}, p, \cP_{out}\}$ is a partition for every propagators $p \in \cP$. Recall from Definition~\ref{admisdfn} that a {\em weakly admissible} diagram is one that satisfies the density and non-crossing conditions but does not require that $|V(\cP)| \geq |\cP| + 4$.

@@

\begin{lem}\label{lem sian}
  Let $W$ be a weakly admissible WLD with at least 5 vertices and in which each vertex supports at least one propagator.  Then at least one of the following two things occurs.
  \begin{enumerate}
    \item $W$ has a propagator of length $\leq 6$ with a propagator of length 2 inside it and nothing else inside it.\label{item big and 2} \sanote{doesn't this depend on which side you call inside? Should we say ``on one side''?  -- the point is that the inside should be the length 6 side.  Now there could be two length 6 sides in which case, yes there should exist one of the length 6 sides with the desired property.  I'm not sure how you want to phrase all this}
    \item There exists a pair of propagators of length $2$ with the property that the first propagator is $(i, i+2)$, the second is $(j, j+2)$, no other propagator ends between vertices $i+2$ and $j+1$, and $j\in\{i+2, i+3, i+4\}$.\label{item pair of 2s}
  \end{enumerate}
\end{lem}

@@

We need a bit of notation:
Define the restriction of $W$ to a region $[i,j]$ to be the subdiagram obtained by forgetting all propagators not wholly supported on vertices in the cyclic interval $[i,j]$; the vertex set remains the same as in $W$.  Denote this diagram by $W_{[i,j]}$. That is, if $W = (\cP, n)$ is a Wilson loop diagram, \bas W_{[i,j]} = (\cP \setminus \Prop([i,j]^c), n)\eas

{\color{blue} If we denote $Q = \Prop([i,j]^c)$, then we can write \bas M(W_{[i,j]}) = M(W)/F(Q) \oplus M_{\emptyset, n-|i-j+1|} \eas where  $M_{\emptyset, k}$ is the matroid of rank $0$ on a set of $k$. That is $M_{\emptyset, n-|i-j+1|} =M((\emptyset, n-|i-j+1|))$. } \sanote{If we like this idea, I can incorporate the necessary nomenclature earlier.}

Note that if $p = (i, j)$, then we can consider two obvious restrictions of $W$, namely \bas W_{[i,j+1]} = W_{V_{in}(p, <_i) \cup V(p)} \quad \textrm{and} \quad  W_{[j,i+1]}  = W_{V_{out}(p, <_i) \cup V(p)} \eas corresponding to the regions inside and outside of $(p,<_i)$.  They contain the propagators on the inside and outside of $(p, <_i)$ respectfully, as well as $p$ itself.  We say that $p$ \textit{bounds the regions} $[i,j+1]$ and $[j,i+1]$ which we call outside and inside respectively. The ordering of the start and end points allow us to identify one or the other region unambiguously.

@@

We will inductively construct a sequence of pairs of propagators $(p_r,q_r)$ satisfying: $\ell(p_r) = 2$, and $p_r$ either forms part of configuration \ref{item big and 2} or \ref{item pair of 2s} from the statement,
  %a configuration \eqref{eq:2 prop config} in $D$
or there is a propagator $q_r$ satisfying
\begin{itemize}
\item $\ell(q_r) \geq 4$.
\item If we orient the $q_r$ such that the $\{p_1 \ldots p_r\}$ are on the outside. The inner region of $W$ that is bounded by $q_r$ is strictly contained in the region of $W$ bounded by $q_{r-1}$ on the inside. Thus the restriction of $W$ to this region does not contain any of the propagators $p_1, \dots, p_r$. \sanote{changed, have I broken it?}
\end{itemize}
Since $W$ contains finitely many vertices, this must eventually terminate in one of the desired configurations. %a configuration of the form \eqref{eq:2 prop config}.

Start by choosing a propagator $p_1 = (j_1,j_1+2)$ of length 2 in $W$ (which exists by Remark~\ref{rem:props of length 2 and 3}).  If it is part of one of the configurations we are looking for then we are done, 
%a configuration \eqref{eq:2 prop config} we are done,
so suppose otherwise. 
By assumption: $p_1$ is not in configuration \ref{item pair of 2s}, there are no propagators of length $3$, and every vertex is supported.  Therefore there must exist a propagator $q_1$ outside of $(p_1, <_j)$ of length $\geq 4$ with one end on edge $j_1$ or on edge $j_1-1$ or on edge $j_1-2$.  This bounds a region $[i_1,k_1] = V_{out}(q_1) \cup V(q_1)$ on the outside of $(q_1, <_{i_1})$ (where $k_1 \in \{j_1-1, j_1, j_1+1\}$) which does not contain $p_1$.

>>>>

This in turn defines two sets of propagators, $\{\cP_{in}(p, <_i), \cP_{out}(p, <_i)\}$, where (for $\bullet \in \{in, \; out\}$) \bas \cP_{\bullet} = \{ q \in (\cP \setminus p) | V(q) \cap V_{\bullet}(p)  \neq \emptyset \} \;.\eas When clear, we drop the parentetical and just write $\cP_{in}$ (resp. $\cP_{out}$). These sets are the sets of propagators, not equal to $p$, \hlfix{whose support contains points in $V_{in}(p, <_*)$ (resp. $V_{in}(p, <_*)$), where $<_*$ refers to the appropriate ordering.}{Note, this means that paralell propagators are neither on the inside and outside of $p$. Is that okay?} If $W$ has no propagators crossing $p$ and no other propagator supported entirely on $V(p)$, then the set $\{\cP_{in}, p, \cP_{out}\}$ partitions $\cP$. Under these circumstances, the set $\cP_{in}$ defines the propagators with end points on one side of $p$, while $\cP_{out}$ defines the propagators with end points on the other side. In particular, for (weakly) admissible Wilson loop diagrams, the set $\{\cP_{in}, p, \cP_{out}\}$ is a partition for every propagators $p \in \cP$. Recall from Definition~\ref{admisdfn} that a {\em weakly admissible} diagram is one that satisfies the density and non-crossing conditions but does not require that $|V(\cP)| \geq |\cP| + 4$.

@@

\begin{lem}\label{lem sian}
  Let $W$ be a weakly admissible WLD with at least 5 vertices and in which each vertex supports at least one propagator.  Then at least one of the following two things occurs.
  \begin{enumerate}
    \item $W$ has a propagator of length $\leq 6$ with a propagator of length 2 on one side it and nothing else on that side.\label{item big and 2} \sanote{tweaked}
    \item There exists a pair of propagators of length $2$ with the property that the first propagator is $(i, i+2)$, the second is $(j, j+2)$, no other propagator ends between vertices $i+2$ and $j+1$, and $j\in\{i+2, i+3, i+4\}$.\label{item pair of 2s}
  \end{enumerate}
\end{lem}

@@

We need a bit of notation: Let $W = (\cP, n)$ a weakly admissible Wilson loop diagram, and $Q \subseteq \cP$. Define the restriction $W|_Q$ to be the diagram $W$ with the propagators not in $Q$ removed. That is , \bas W|_Q = (Q, n). \eas  \footnote{If we denote $V \subset [n]$ and $Q = \Prop(V^c)$, then we can write \bas M(W|_{Q^c}) = M(W)/F(Q) \oplus M_{\emptyset, n-|i-j+1|} \eas where  $M_{\emptyset, k}$ is the matroid of rank $0$ on a set of $k$. That is $M_{\emptyset, n-|i-j+1|} =M((\emptyset, n-|i-j+1|))$. } \sanote{tweaked this para and the next}

%to a region $[i,j]$ to be the subdiagram obtained by forgetting all propagators not wholly supported on vertices in the cyclic interval $[i,j]$; the vertex set remains the same as in $W$.  Denote this diagram by $W_{[i,j]}$. That is, if $W = (\cP, n)$ is a Wilson loop diagram, \bas W_{[i,j]} = (\cP \setminus \Prop([i,j]^c), n)\eas

Note that if $p = (i, j)$, then we can consider two obvious restrictions of $W$, namely \bas W|_{\cP_{in} \cup p} \quad \textrm{and} \quad  W|_{\cP_{out}\cup p}, \eas where the set of non-supportive vertices of  $W|_{\cP_{in} \cup p}$ is $V_{out}$ and  the set of non-supportive vertices of  $W|_{\cP_{out} \cup p}$ is $V_{in}$. That is, these restrictions contain the propagators on the inside and outside of $(p, <_i)$ respectfully, as well as $p$ itself.

%We say that $p$ \textit{bounds the regions} $[i,j+1]$ and $[j,i+1]$ which we call outside and inside respectively. The ordering of the start and end points allow us to identify one or the other region unambiguously.

@@

We will inductively construct a sequence of pairs of propagators $(p_r,q_r)$ satisfying: $\ell(p_r) = 2$, and $p_r$ either forms part of configuration \ref{item big and 2} or \ref{item pair of 2s} from the statement,
  %a configuration \eqref{eq:2 prop config} in $D$
or there is a propagator $q_r$ satisfying
\begin{itemize}
\item $\ell(q_r) \geq 4$.
\item If we orient the $q_r = (i_1, k_1)$ such that the $\{p_1 \ldots p_r\} \subseteq \cP_{out}(q_r)$ are on the outside. Furthermore, we have have $\cP_{in}(q_r) \subsetneq \cP_{in}(q_{r-1})$. That is, the propagators on the inside of $q_r$ are strictly contained in the set of propagators inside of $q_{r-1}$. In particular, $W_{\cP_{in}(q_r) \cup q_r}$ does not contain any of the propagators $p_1, \dots, p_r$. \sanote{changed, have I broken it?}
\end{itemize}
Since $W$ contains finitely many vertices, this must eventually terminate in one of the desired configurations. %a configuration of the form \eqref{eq:2 prop config}.

Start by choosing a propagator $p_1 = (j_1,j_1+2)$ of length 2 in $W$ (which exists by Remark~\ref{rem:props of length 2 and 3}).  If it is part of one of the configurations we are looking for then we are done, 
%a configuration \eqref{eq:2 prop config} we are done,
so suppose otherwise. 
By assumption: $p_1$ is not in configuration \ref{item pair of 2s}, there are no propagators of length $3$, and every vertex is supported.  Therefore there must exist a propagator $q_1$ of length $\geq 4$ with one end on edge $j_1$ or on edge $j_1-1$ or on edge $j_1-2$ (that is, $k_1 \in \{j_1-1, j_1, j_1+1\}$). Orient $q_1$ such that  $p_1 \in \cP_{out}(q_1)$, (that is $(q_1, <_{i_1})$).

%This bounds a region $[i_1,k_1] = V_{out}(q_1) \cup V(q_1)$ on the outside of $(q_1, <_{i_1})$ (where $k_1 \in \{j_1-1, j_1, j_1+1\}$) which does not contain $p_1$.
